---
title: "Petroeum oil"
author: "Efoli Matthew"
date: "2/19/2021"
output: html_document
---

# Objective of the data visualization
Make a bar plot with SN on the y-axis, compounds on the x-axis, SD as the error bars.  
separating the isomers by colours for each distinctive compounds.
Expand the y-axis to about 350. 

Note: The chart is flipped.


## Import data

The data set has been cleaned in Microsoft Excel. 
A tip: It is easier to clean your data from the source to which it is from. Although, I would still perform some basic check with my  favourite snippets of codes.

```{r import data, echo=FALSE}
#import data
Barplot2 <- read.csv("C:/Users/test/Documents/Github projects/petroleumoil/Barplotinput2.csv", header = T)
# structure of data
str(Barplot2)
Barplot3 <- Barplot2[, 1:10] # doing away with the last 3 columns.
str(Barplot3)
# check for missing values
sum(is.na(Barplot3))
Barplot3[,10]
typeof(Barplot3[ ,10])
```

## Adding the distinct colouring

The colours  were added to distinguish the chemical compounds.

```{r addcolours}
cols <- c("C10H8" = "darkorange2", "C11H10" = "dodgerblue2", "C12H8" = "firebrick2", "C12H10" = "darkgreen", "C12H12" = "turquoise3", "C13H10" = "gray", "C13H12" = "violet", "C13H14" = "yellowgreen", "C14H10" = "blue4", "C14H12" = "red", "C14H14" = "peru", "C14H16" = "darkolivegreen4", "C12H12N2" = "slateblue", "C13H12O" = "deeppink4", "C15H12" = "chocolate", "C15H14" = "steelblue2", "C15H16" = "grey50", "C13H10S" = "yellow", "C16H10" = "hotpink", "C16H12" = "navy", "C16H14" = "maroon", "C17H14" = "green1", "C18H12" = "lightcoral", "C18H18" = "slateblue1", "C17H14O" = "magenta1", "C20H12" = "goldenrod2", "C22H12" = "springgreen4", "C22H14" = "black")
```

### load necessary packages
```{r loadpackages}
library(ggplot2)
library(ggpubr)
```


## The visualization

```{r Calvert, echo=FALSE}
Calvert <- ggplot(Barplot3, aes(x=S.N, y= Calvert, fill= Compound)) + geom_bar(stat = "identity", position = position_dodge(0.9))+ geom_errorbar(aes(ymin=Calvert-sd1, ymax=Calvert+sd1), width=0.15, size=0.5, position = position_dodge(0.9),alpha=0.3) + coord_flip() + theme(axis.title.y = element_blank(), legend.position = "none") + scale_fill_manual(values = cols) + expand_limits(y = c(1, 350))
Calvert
```

```{r Macondo}
Macondo <- ggplot(Barplot3, aes(x=S.N, y= Macondo, fill= Compound)) + geom_bar(stat = "identity", position = position_dodge(0.9))+ geom_errorbar(aes(ymin=Macondo-sd1, ymax=Macondo+sd1), width=0.15, size=0.5, position = position_dodge(0.9),alpha=0.3) + coord_flip() + theme(axis.title.y = element_blank(), legend.position = "none") + scale_fill_manual(values = cols) + expand_limits(y = c(1, 350)) 

Macondo
```

```{r Duri}
Duri <- ggplot(Barplot3, aes(x=S.N, y= Duri, fill= Compound)) + geom_bar(stat = "identity", position = position_dodge(0.9))+ geom_errorbar(aes(ymin=Duri-sd1, ymax=Duri+sd1), width=0.15, size=0.5, position = position_dodge(0.9),alpha=0.3) + coord_flip() + theme(axis.title.y = element_blank(), legend.position = "none") + scale_fill_manual(values = cols) + expand_limits(y = c(1, 350))
Duri
```


```{r SanArdo}
SanArdo <- ggplot(Barplot3, aes(x=S.N, y= SanArdo, fill= Compound)) + geom_bar(stat = "identity", position = position_dodge(0.9))+ geom_errorbar(aes(ymin=SanArdo-sd1, ymax=SanArdo+sd1), width=0.15, size=0.5, position = position_dodge(0.9),alpha=0.3) + coord_flip() + theme(axis.title.y = element_blank()) + scale_fill_manual(values = cols) + expand_limits(y = c(1, 350))
SanArdo
```

```{r themes}
mytheme <- theme(panel.grid = element_blank(),panel.background = element_rect(fill = "white"),axis.text.y = element_blank(), axis.ticks.y.left = element_blank(), axis.line.x = element_line(colour = "black", size = 0.2),axis.line.y = element_line(colour = "black", size = 0.2))

mytheme2 <- theme(panel.grid = element_blank(),panel.background = element_rect(fill = "white"),axis.text.y = element_blank(), axis.ticks.y.left = element_blank(), axis.line.x = element_line(colour = "black", size = 0.2)) #this theme has no y-axis note.
```

```{r EPAdataviz2}
par(mfrow=c(1,4)) #put the code in a 1 row, 4 columns format
Calvert + mytheme
Macondo + mytheme2
Duri + mytheme2
SanArdo + mytheme2

Ca <- Calvert + mytheme
Mac <- Macondo + mytheme2
Dur <- Duri + mytheme2
San <- SanArdo + mytheme2
ggarrange(Ca, Mac, Dur, San, ncol = 4, nrow = 1)


ggsave("EDA7.pdf", width=20, height = 10)  #save the code in a PDF file 
ggsave("EDA8.TIFF", width=30, height = 25) #save the code as a TIFF file
```

